---
title: 'S1 Text: Modelling results'
author: 'Comparative analysis of dengue and Zika outbreaks reveals differences by setting and virus'
date: 'Sebastian Funk, Adam J. Kucharski, Anton Camacho, Rosalind M. Eggo, Laith Yakob, Lawrence Murray, W. John Edmunds'
output:
  html_document:
    self_contained: true
    toc: false
    toc_depth: 3
---

```{r setup, echo = FALSE, cache = FALSE}
knitr::opts_chunk$set(cache = TRUE, fig.cap = '', fig.align = "center", message = FALSE, tidy = TRUE, warning = FALSE)
```

# Introduction

This document contains the R code necessary to reproduce the results in "Comparative analysis of dengue and Zika outbreaks reveals differences by setting and virus." All code and data used here are in a [github repository](http://github.com/sbfnk/vbd). The data sets for the three outbreaks are contained in the `data` subdirectory.

## Required packages
```{r packages}
cran_packages <- c("cowplot", "dplyr", "tidyr", "stringi", "msm", "rbi")
github_packages <- c("sbfnk/RBi.helpers")
```

```{r install, eval = FALSE}
for (package in cran_packages)
{
    install.packages(package)
}

library('devtools')

for (package in github_packages)
{
    install_github(package)
}
```

# Generate MCMC chains

Generating MCMC chains requires a working installation of [libbi](http://libbi.org/) v1.2.0. The model code is in [vbd.bi](https://github.com/sbfnk/vbd/blob/master/bi/vbd.bi), which is run by the R script [dengue_zika_mcmc.r](https://github.com/sbfnk/vbd/blob/master/R/dengue_zika_mcmc.r).

Edit the lines starting `code_dir <- ` and `output_dir <-` in the script 'dengue_zika_mcmc.r' from the R/ directory in the github repository to point to the correct directories, and run it using

```{sh eval = FALSE}
Rscript dengue_zika_mcmc.r --nsamples 10000000 --pre_samples 10000 --sample-prior --sample-observations --thin 1000 --fix-natural-history
Rscript dengue_zika_mcmc.r --nsamples 10000000 --pre_samples 10000 --sample-prior --sample-observations --thin 1000 --fix-natural-history --shorter-mosquito-lifespan
```

The first line runs with $D_\mathrm{life,M}=2\,\mathrm{weeks}$, and the second one with $D_\mathrm{life,M}=1\,\mathrm{week}$


# Analyse MCMC chains

Set the `code_dir` and `output_dir` to the value used in `sample.posterior.r`

```{r eval = FALSE}
output_dir <- 'insert your directory here'
code_dir <- 'insert your directory here'
```

```{r echo = FALSE}
output_dir <- "~/Data/Zika/"
code_dir <- "~/code/vbd/"
```

Load required packages
```{r load, cache = FALSE}
for (package in c(cran_packages, github_packages))
{
    library(tolower(sub("^.*/", "", package)), character.only = TRUE)
}
```

Load script from github repository (add path if necessary)

```{r load_script, cache = FALSE}
source(paste0(code_dir, "/R/", "analyse_traces.r"))
```

Get MCMC chains of the two libbi runs.

```{r get_chains}
libbi_results <- analyse_traces(c("vbd_fnh", "vbd_fnh_shorter"), output_dir)
``` 

Calculate DIC for models with different mosquito life times (1 week vs. 2 weeks vs. both with equal probability)

```{r DIC}
dic <- sapply(libbi_results, function(x) {compute_DIC(x[["trace"]][["posterior"]])})
dic
```

# Plots

## Prepare data

```{r prepare_data}
ts <- list()
analyses <- data.frame(setting = c("yap", "yap", "fais"), disease = c("dengue", "zika", "dengue"))

for (i in 1:nrow(analyses))
{
    this_setting <- analyses[i, "setting"]
    this_disease <- analyses[i, "disease"]
    this_filename <-
      paste(code_dir, "data",
            paste(this_setting, this_disease, "data.rds", sep = "_"),
            sep = "/")
    this_ts <- readRDS(this_filename) %>%
      mutate(setting = this_setting, disease = this_disease,
             week = floor(nr / 7))
    ts <- c(ts, list(this_ts))
}

ordered_obs_id_levels <- c("yap_zika", "yap_dengue", "fais_dengue")

data_labels <- ordered_obs_id_levels
data_labels <- sub("^(.*)_(.*)$", "\\2 \\1", data_labels)
data_labels <- sub(" ", " in ", stri_trans_totitle(data_labels))
names(data_labels) <- ordered_obs_id_levels

data <- bind_rows(ts) %>%
    group_by(week, setting, disease) %>%
    summarize(value = sum(value), onset_date = min(onset_date)) %>%
    ungroup() %>%
    mutate(obs_id = factor(paste(setting, disease, sep = "_"),
                           levels = ordered_obs_id_levels,
                           labels = data_labels)) %>%
    arrange(week, obs_id) %>%
    select(week, obs_id, value, onset_date) %>% 
    rename(time = week) %>%
    mutate(state = "Cases")
first_obs <- data %>%
    group_by(obs_id) %>%
    filter(value > 0) %>%
    slice(which.min(time)) %>%
    select(time, obs_id) %>%
    rename(first_obs = time)
last_obs <- data %>%
    group_by(obs_id) %>%
    filter(value > 0) %>%
    slice(which.max(time)) %>%
    select(time, obs_id) %>%
    rename(last_obs = time)
data <- data %>%
    left_join(first_obs, by = "obs_id") %>%
    left_join(last_obs, by = "obs_id") %>%
    filter(time >= first_obs & time <= last_obs)

``` 

## Fits


```{r plot_fits, fig.height = 3}
  libbi_results[["vbd_fnh_all"]][["trace"]][["posterior"]][["Cases"]] <- 
    libbi_results[["vbd_fnh_all"]][["trace"]][["posterior"]][["Cases"]] %>% 
      mutate(obs_id = factor(obs_id, levels = ordered_obs_id_levels,
                                     labels = data_labels))
  temp_plot <-
    plot_libbi(read = libbi_results[["vbd_fnh_all"]][["trace"]][["posterior"]]["Cases"],
               model = libbi_results[["vbd_fnh_all"]][["model"]],
               data = data %>% filter(value > 0),
               density_args = list(adjust = 2),
               extra.aes = list(group = "obs_id"),
               data.colour = "black",
               states = "Cases", trend = "mean", plot = FALSE,
               limit.to.data = TRUE,
               quantiles = c(0.5, 0.72, 0.95))
  obs_states <- temp_plot$data$states %>%
      inner_join(data %>% select(time, obs_id, onset_date), by = c("time", "obs_id"))
  p_obs <- ggplot(obs_states, aes(x = onset_date)) +
      geom_point(data = data, mapping = aes(y = value)) +
      facet_wrap(~ obs_id, scales = "free") +
      scale_x_date("Week", labels = scales::date_format("%e %b %Y")) +
      theme_cowplot(font_size = 12) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      facet_wrap(~ obs_id, scales = "free") +
      scale_y_continuous("Disease incidence") +
      geom_line(aes(y = value)) +
      geom_ribbon(aes(ymin = min.1, ymax = max.1), alpha = 0.5) +
      geom_ribbon(aes(ymin = min.2, ymax = max.2), alpha = 0.25) +
      geom_ribbon(aes(ymin = min.3, ymax = max.3), alpha = 0.125)

  p_obs
```

## Other plots

Define plot labels

```{r labels}
labels <- c(p_d_inc_h = "italic(D)[plain(inc,H)]",
            p_d_inc_m = "italic(D)[plain(inc,M)]",
            p_d_inf_h = "italic(D)[plain(inf,H)]",
            p_lm = "log[10](italic(m))",
            p_initial_susceptible_yap = "italic(q)",
            p_rep = "italic(r)",
            p_b_h = "italic(b)[H]",
            p_b_m = "italic(b)[M]",
            p_t_start = "italic(t[0])",
            p_pop_yap = "italic(rho)",
            p_red_foi_yap = "italic(sigma)",
            p_p_patch_yap = "italic(phi)",
            R0 = "italic(R)[H %->% H]",
            GI = "italic(G)",
            zika = "Zika",
            yap = "Yap",
            fais = "Fais")

```

### Marginal posterior densities

```{r plot_libbi, cache = FALSE, fig.height = 8}
      p <-
          plot_libbi(read = libbi_results[["vbd_fnh_all"]][["trace"]][["posterior"]],
                     prior = libbi_results[["vbd_fnh_all"]][["trace"]][["prior"]],
                     model = libbi_results[["vbd_fnh_all"]][["model"]],
                     density_args = list(bins = 20, alpha = 0.5, color = "black"),
                     densities = "histogram",
                     extra.aes = list(color = "disease", linetype = "setting"),
                     trend = "median", plot = FALSE,
                     quantiles = c(0.5, 0.95),
                     labels = labels, brewer.palette = "Set1")
     p$densities
```

### Human-to-human reproduction number vs generation interval

```{r r0vgi, cache = FALSE,fig.height = 3}
temp_plot <-
    plot_libbi(read = libbi_results[["vbd_fnh_all"]][["trace"]][["posterior"]],
               prior = libbi_results[["vbd_fnh_all"]][["trace"]][["prior"]],
               model = libbi_results[["vbd_fnh_all"]][["model"]],
               extra.aes = list(color = "disease",
                                linetype = "setting",
                                group = "p_d_life_m"),
               plot = FALSE,
               labels = labels,
               states = c(),
               params = c("R0", "GI"),
               noises = c())
               
r0gi <- temp_plot$data$params %>%
    filter(distribution == "posterior") %>%
    mutate(obs_id = tolower(paste(setting, disease, sep = "_"))) %>%
    filter(obs_id != "fais_zika") %>%
    mutate(obs_id = factor(obs_id, levels = ordered_obs_id_levels,
                               labels = data_labels)) %>%
    spread(parameter, value)

cross_sections <- data.frame(GI = c(3, 4))
colnames(cross_sections) <- "italic(G)"

p_r0vgi <- ggplot(r0gi, aes(x = `italic(G)` * 7)) +
        geom_jitter(aes(y = `italic(R)[H %->% H]`, color = factor(p_d_life_m))) +
        facet_grid(~ obs_id) +
        scale_x_continuous("Equilibrium generation interval (days)") +
        scale_y_continuous(expression(italic(R)[H %->% H])) +
        scale_color_brewer("Mosquito life span", palette = "Dark2", labels = c("1 week", "2 weeks")) +
        theme_cowplot(font_size = 12) +
        theme(legend.position = "top") + 
        facet_wrap(~ obs_id, scales = "free") +
        geom_rect(data = cross_sections, aes(xmin = `italic(G)` * 7 - 1, xmax = `italic(G)` * 7 + 1), ymin = -Inf, ymax = Inf, alpha = 0.2, lwd = 3)
p_r0vgi
ggsave("r0vgi.pdf", p_r0vgi, height = 4, width = 8)

quantile_probs <- c(0.5, 0.25, 0.75, 0.025, 0.975)

r0gi %>% 
  filter(abs(`italic(G)` - round(`italic(G)`)) <= 1/7) %>% 
  mutate(GI = round(`italic(G)`)) %>% 
  filter(GI %in% c(3, 4)) %>% 
  group_by(disease, setting, GI) %>% 
  do(data.frame(prob = factor(quantile_probs, levels = quantile_probs), 
                value = quantile(.$`italic(R)[H %->% H]`, quantile_probs))) %>% 
  spread(prob, value)
```

### Traces

```{r traces, fig.height = 10}
p$traces
```

# Other models 

These are the alternative models considered and mentioned in the manuscript.

## Reduced population size in Yap
### Generate MCMC chains

In this model, the population in Yap is reduced by a factor $rho$.

Fit the model (remember to update `code_dir` and `data_dir` in `dengue_zika_mcmc.r`):

```{sh eval = FALSE}
Rscript dengue_zika_mcmc.r --nsamples 10000000 --pre_samples 10000 --sample-prior --sample-observations --thin 1000 --fix-natural-history --sero --pop
Rscript dengue_zika_mcmc.r --nsamples 10000000 --pre_samples 10000 --sample-prior --sample-observations --thin 1000 --fix-natural-history --shorter-mosquito-lifespan --sero --pop
```

Get MCMC chains of the two libbi runs.

```{r get_chains_red_pop}
libbi_results <- analyse_traces(c("vbd_sero_pop_fnh", "vbd_sero_pop_fnh_shorter"), output_dir)
``` 

### Fits

```{r plot_fits_red_pop}
  libbi_results[["vbd_sero_pop_fnh_all"]][["trace"]][["posterior"]][["Cases"]] <- 
    libbi_results[["vbd_sero_pop_fnh_all"]][["trace"]][["posterior"]][["Cases"]] %>% 
      mutate(obs_id = factor(obs_id, levels = ordered_obs_id_levels,
                                     labels = data_labels))
   temp_plot <-
    plot_libbi(read = libbi_results[["vbd_sero_pop_fnh_all"]][["trace"]][["posterior"]]["Cases"],
               model = libbi_results[["vbd_sero_pop_fnh_all"]][["model"]],
               data = data %>% filter(value > 0),
               density_args = list(adjust = 2),
               extra.aes = list(group = "obs_id"),
               data.colour = "black",
               states = "Cases", trend = "mean", plot = FALSE,
               limit.to.data = TRUE,
               quantiles = c(0.5, 0.72, 0.95))
  obs_states <- temp_plot$data$states %>%
      inner_join(data %>% select(time, obs_id, onset_date), by = c("time", "obs_id"))
  p_obs <- ggplot(obs_states, aes(x = onset_date)) +
      geom_point(data = data, mapping = aes(y = value)) +
      facet_wrap(~ obs_id, scales = "free") +
      scale_x_date("Week", labels = scales::date_format("%e %b %Y")) +
      theme_cowplot(font_size = 12) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      facet_wrap(~ obs_id, scales = "free") +
      scale_y_continuous("Disease incidence") +
      geom_line(aes(y = value)) +
      geom_ribbon(aes(ymin = min.1, ymax = max.1), alpha = 0.5) +
      geom_ribbon(aes(ymin = min.2, ymax = max.2), alpha = 0.25) +
      geom_ribbon(aes(ymin = min.3, ymax = max.3), alpha = 0.125)

  p_obs
```

### Marginal posterior densities

```{r plot_libbi_red_pop, cache = FALSE, fig.height = 8}
      p <-
          plot_libbi(read = libbi_results[["vbd_sero_pop_fnh_all"]][["trace"]][["posterior"]],
                     prior = libbi_results[["vbd_sero_pop_fnh_all"]][["trace"]][["prior"]],
                     model = libbi_results[["vbd_sero_pop_fnh_all"]][["model"]],
                     density_args = list(bins = 20, alpha = 0.5, color = "black"),
                     densities = "histogram",
                     extra.aes = list(color = "disease", linetype = "setting"),
                     trend = "median", plot = FALSE,
                     quantiles = c(0.5, 0.95),
                     labels = labels, brewer.palette = "Set1")
     p$densities
```

### Estimated final size for Zika in Yap (proportion infected in the outbreak)

```{r final_size_pop}
n_infected <- libbi_results[["vbd_sero_pop_fnh_all"]][["trace"]][["posterior"]][["final_size"]] %>% filter(obs_id == "yap_zika") %>% .$value %>% quantile(c(0.025, 0.5, 0.975)) 
N <- libbi_results[["vbd_sero_pop_fnh_all"]][["trace"]][["posterior"]][["p_N_h"]] %>% filter(setting == "yap") %>% .$value %>% mean
n_infected / N
```

## Reduced susceptibility against Zika in Yap
### Generate MCMC chains

In this model, the population susceptible to Zika in Yap is reduced by a factor $r$.

Fit the model (remember to update `code_dir` and `data_dir` in `dengue_zika_mcmc.r`):

```{sh eval = FALSE}
Rscript dengue_zika_mcmc.r --nsamples 10000000 --pre_samples 10000 --sample-prior --sample-observations --thin 1000 --fix-natural-history --sero 
Rscript dengue_zika_mcmc.r --nsamples 10000000 --pre_samples 10000 --sample-prior --sample-observations --thin 1000 --fix-natural-history --shorter-mosquito-lifespan --sero
```

Get MCMC chains of the two libbi runs

```{r get_chains_red_susc}
libbi_results <- analyse_traces(c("vbd_sero_fnh", "vbd_sero_fnh_shorter"), output_dir)
``` 

### Fits

```{r plot_fits_red_susc}
  libbi_results[["vbd_sero_fnh_all"]][["trace"]][["posterior"]][["Cases"]] <- 
    libbi_results[["vbd_sero_fnh_all"]][["trace"]][["posterior"]][["Cases"]] %>% 
      mutate(obs_id = factor(obs_id, levels = ordered_obs_id_levels,
                                     labels = data_labels))
   temp_plot <-
    plot_libbi(read = libbi_results[["vbd_sero_fnh_all"]][["trace"]][["posterior"]]["Cases"],
               model = libbi_results[["vbd_sero_fnh_all"]][["model"]],
               data = data %>% filter(value > 0),
               density_args = list(adjust = 2),
               extra.aes = list(group = "obs_id"),
               data.colour = "black",
               states = "Cases", trend = "mean", plot = FALSE,
               limit.to.data = TRUE,
               quantiles = c(0.5, 0.72, 0.95))
  obs_states <- temp_plot$data$states %>%
      inner_join(data %>% select(time, obs_id, onset_date), by = c("time", "obs_id"))
  p_obs <- ggplot(obs_states, aes(x = onset_date)) +
      geom_point(data = data, mapping = aes(y = value)) +
      facet_wrap(~ obs_id, scales = "free") +
      scale_x_date("Week", labels = scales::date_format("%e %b %Y")) +
      theme_cowplot(font_size = 12) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      facet_wrap(~ obs_id, scales = "free") +
      scale_y_continuous("Disease incidence") +
      geom_line(aes(y = value)) +
      geom_ribbon(aes(ymin = min.1, ymax = max.1), alpha = 0.5) +
      geom_ribbon(aes(ymin = min.2, ymax = max.2), alpha = 0.25) +
      geom_ribbon(aes(ymin = min.3, ymax = max.3), alpha = 0.125)

  p_obs
```

### Marginal posterior densities

```{r plot_libbi_red_susc, cache = FALSE, fig.height = 8}
      p <-
          plot_libbi(read = libbi_results[["vbd_sero_fnh_all"]][["trace"]][["posterior"]],
                     prior = libbi_results[["vbd_sero_fnh_all"]][["trace"]][["prior"]],
                     model = libbi_results[["vbd_sero_fnh_all"]][["model"]],
                     density_args = list(bins = 20, alpha = 0.5, color = "black"),
                     densities = "histogram",
                     extra.aes = list(color = "disease", linetype = "setting"),
                     trend = "median", plot = FALSE,
                     quantiles = c(0.5, 0.95),
                     labels = labels, brewer.palette = "Set1")
     p$densities
```

### Estimated final size for Zika in Yap (proportion infected in the outbreak)

```{r final_size_sero}
n_infected <- libbi_results[["vbd_sero_fnh_all"]][["trace"]][["posterior"]][["final_size"]] %>% filter(obs_id == "yap_zika") %>% .$value %>% quantile(c(0.025, 0.5, 0.975)) 
N <- libbi_results[["vbd_sero_fnh_all"]][["trace"]][["posterior"]][["p_N_h"]] %>% filter(setting == "yap") %>% .$value %>% mean
n_infected / N
```

## Two-patch model

Fit the model (remember to update `code_dir` and `data_dir` in `dengue_zika_mcmc.r`):

```{sh eval = FALSE}
Rscript dengue_zika_mcmc.r --nsamples 10000000 --pre_samples 10000 --sample-prior --sample-observations --thin 1000 --fix-natural-history --patch
Rscript dengue_zika_mcmc.r --nsamples 10000000 --pre_samples 10000 --sample-prior --sample-observations --thin 1000 --fix-natural-history --shorter-mosquito-lifespan --patch
```

Get MCMC chains of the two libbi runs.

```{r get_chains_two_patch}
libbi_results <- analyse_traces(c("vbd_patch_fnh", "vbd_patch_fnh_shorter"), output_dir)
``` 

### Fits

```{r plot_fits_two_patch}
  libbi_results[["vbd_patch_fnh_all"]][["trace"]][["posterior"]][["Cases"]] <- 
    libbi_results[["vbd_patch_fnh_all"]][["trace"]][["posterior"]][["Cases"]] %>% 
      mutate(obs_id = factor(obs_id, levels = ordered_obs_id_levels,
                                     labels = data_labels))
   temp_plot <-
    plot_libbi(read = libbi_results[["vbd_patch_fnh_all"]][["trace"]][["posterior"]]["Cases"],
               model = libbi_results[["vbd_patch_fnh_all"]][["model"]],
               data = data %>% filter(value > 0),
               density_args = list(adjust = 2),
               extra.aes = list(group = "obs_id"),
               data.colour = "black",
               states = "Cases", trend = "mean", plot = FALSE,
               limit.to.data = TRUE,
               quantiles = c(0.5, 0.72, 0.95))
  obs_states <- temp_plot$data$states %>%
      inner_join(data %>% select(time, obs_id, onset_date), by = c("time", "obs_id"))
  p_obs <- ggplot(obs_states, aes(x = onset_date)) +
      geom_point(data = data, mapping = aes(y = value)) +
      facet_wrap(~ obs_id, scales = "free") +
      scale_x_date("Week", labels = scales::date_format("%e %b %Y")) +
      theme_cowplot(font_size = 12) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      facet_wrap(~ obs_id, scales = "free") +
      scale_y_continuous("Disease incidence") +
      geom_line(aes(y = value)) +
      geom_ribbon(aes(ymin = min.1, ymax = max.1), alpha = 0.5) +
      geom_ribbon(aes(ymin = min.2, ymax = max.2), alpha = 0.25) +
      geom_ribbon(aes(ymin = min.3, ymax = max.3), alpha = 0.125)

  p_obs
```

### Marginal posterior densities

```{r plot_libbi_two_patch, cache = FALSE, fig.height = 8}
      p <-
          plot_libbi(read = libbi_results[["vbd_patch_fnh_all"]][["trace"]][["posterior"]],
                     prior = libbi_results[["vbd_patch_fnh_all"]][["trace"]][["prior"]],
                     model = libbi_results[["vbd_patch_fnh_all"]][["model"]],
                     density_args = list(bins = 20, alpha = 0.5, color = "black"),
                     densities = "histogram",
                     extra.aes = list(color = "disease", linetype = "setting"),
                     trend = "median", plot = FALSE,
                     quantiles = c(0.5, 0.95),
                     labels = labels, brewer.palette = "Set1")
     p$densities
```

### Fitting the Zika outbreak in Yap in isolation

Fit the model (remember to update `code_dir` and `data_dir` in `dengue_zika_mcmc.r`):

```{sh eval = FALSE}
Rscript dengue_zika_mcmc.r --nsamples 10000000 --pre_samples 10000 --sample-prior --sample-observations --thin 1000 --fix-natural-history --patch --disease zika --setting yap
Rscript dengue_zika_mcmc.r --nsamples 10000000 --pre_samples 10000 --sample-prior --sample-observations --thin 1000 --fix-natural-history --shorter-mosquito-lifespan --patch --disease zika --setting yap
```

Get MCMC chains of the two libbi runs.

```{r get_chains_two_patch_zika_yap}
libbi_results <- analyse_traces(c("vbd_patch_fnh_yap_zika", "vbd_patch_fnh_shorter_yap_zika"), output_dir)
``` 

Fit

```{r plot_fits_two_patch_zika_yap}
  libbi_results[["vbd_patch_fnh_yap_zika_all"]][["trace"]][["posterior"]][["Cases"]] <- 
    libbi_results[["vbd_patch_fnh_yap_zika_all"]][["trace"]][["posterior"]][["Cases"]] %>% 
      mutate(obs_id = factor(obs_id, levels = ordered_obs_id_levels,
                                     labels = data_labels))
   temp_plot <-
    plot_libbi(read = libbi_results[["vbd_patch_fnh_yap_zika_all"]][["trace"]][["posterior"]]["Cases"],
               model = libbi_results[["vbd_patch_fnh_yap_zika_all"]][["model"]],
               data = data,
               density_args = list(adjust = 2),
               extra.aes = list(group = "obs_id"),
               data.colour = "black",
               states = "Cases", trend = "mean", plot = FALSE,
               limit.to.data = TRUE,
               quantiles = c(0.5, 0.72, 0.95))
  obs_states <- temp_plot$data$states %>%
      inner_join(data %>% select(time, obs_id, onset_date),
		 by = c("time", "obs_id"))
  
  p_obs <- ggplot(obs_states, aes(x = onset_date)) +
      geom_point(data = data %>% filter(obs_id == "Zika in Yap"), mapping = aes(y = value)) +
      facet_wrap(~ obs_id, scales = "free") +
      scale_x_date("Week", labels = scales::date_format("%e %b %Y")) +
      theme_cowplot(font_size = 12) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      facet_wrap(~ obs_id, scales = "free") +
      scale_y_continuous("Disease incidence") +
      geom_line(aes(y = value)) +
      geom_ribbon(aes(ymin = min.1, ymax = max.1), alpha = 0.5) +
      geom_ribbon(aes(ymin = min.2, ymax = max.2), alpha = 0.25) +
      geom_ribbon(aes(ymin = min.3, ymax = max.3), alpha = 0.125)

  p_obs
```

Marginal posterior densities

```{r plot_libbi_two_patch_zika_yap, cache = FALSE, fig.height = 8}
      p <-
          plot_libbi(read = libbi_results[["vbd_patch_fnh_yap_zika_all"]][["trace"]][["posterior"]],
                     prior = libbi_results[["vbd_patch_fnh_yap_zika_all"]][["trace"]][["prior"]],
                     model = libbi_results[["vbd_patch_fnh_yap_zika_all"]][["model"]],
                     density_args = list(bins = 20, alpha = 0.5, color = "black"),
                     densities = "histogram",
                     extra.aes = list(color = "disease", linetype = "setting"),
                     trend = "median", plot = FALSE,
                     quantiles = c(0.5, 0.95),
                     labels = labels, brewer.palette = "Set1")
     p$densities
```

Human-to-human reproduction number vs generation interval

```{r r0vgi_patch_zika_yap, cache = FALSE,fig.height = 3}
temp_plot <-
    plot_libbi(read = libbi_results[["vbd_patch_fnh_yap_zika_all"]][["trace"]][["posterior"]],
               prior = libbi_results[["vbd_patch_fnh_yap_zika_all"]][["trace"]][["prior"]],
               model = libbi_results[["vbd_patch_fnh_yap_zika_all"]][["model"]],
               extra.aes = list(color = "disease",
                                linetype = "setting",
                                group = "p_d_life_m"),
               plot = FALSE,
               labels = labels,
               states = c(),
               params = c("R0", "GI"),
               noises = c())
               
r0gi <- temp_plot$data$params %>%
    filter(distribution == "posterior") %>%
    mutate(obs_id = tolower(paste(setting, disease, sep = "_"))) %>%
    filter(obs_id != "fais_zika") %>%
    mutate(obs_id = factor(obs_id, levels = ordered_obs_id_levels,
                               labels = data_labels)) %>%
    spread(parameter, value)

p_r0vgi <- ggplot(r0gi, aes(x = `italic(G)` * 7)) +
        geom_jitter(aes(y = `italic(R)[H %->% H]`, color = factor(p_d_life_m))) +
        facet_grid(~ obs_id) +
        scale_x_continuous("Equilibrium generation interval (days)") +
        scale_y_continuous(expression(italic(R)[H %->% H])) +
        scale_color_brewer("Mosquito life span", palette = "Dark2", labels = c("1 week", "2 weeks")) +
        theme_cowplot(font_size = 12) +
        theme(legend.position = "top") + 
        facet_wrap(~ obs_id, scales = "free") +
        geom_rect(data = cross_sections, aes(xmin = `italic(G)` * 7 - 1, xmax = `italic(G)` * 7 + 1), ymin = -Inf, ymax = Inf, alpha = 0.2, lwd = 3)
p_r0vgi
```
